---
title: "Bridgeport + Stamford extended"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(tidycensus)
```

```{r}
################# VARIABLES #######
year <- 2016
stam <- read_csv("stamford_tract_shares.csv")
bpt <- read_csv("bpt_tract_shares.csv") %>% rename(share = reshare)
geo <- bind_rows(stam %>% mutate(city = "Stamford"), bpt %>% mutate(city = "Bridgeport")) %>%
	mutate(name = str_replace_all(name, "_", " ") %>% str_to_title())
# clean up neighborhood names later
tracts <- geo$trct
city_names <- c("Bridgeport", "Stamford")
county <- "09001"
#####################################

acs_vars <- read_csv("../reference/acs_variables.csv")
```

```{r}
make_share <- function(df, denom) {
	df %>%
		mutate(share = ifelse(group == denom, NA, round(value / value[group == denom], digits = 3)))
}

add_grp <- function(df, ...) {
	cols <- quos(...)
	select(df, !!!cols) %>% reduce(add)
}
```

```{r}
table_nums <- list(
  total_pop = "B01003",
  sex_age = "B01001",
  race = "B03002",
  family = "B11001",
  children = "B11003",
  vacancy = "B25004",
  tenure = "B25003",
  disconnect = "B14005",
  education = "B15002",
  immigration = "B05005",
  language2 = "B16004",
  labor = "B23025",
  commute = "B08301",
  parent_work = "B23008",
  occupation = "C24010",
  hh_income = "B19001",
  agg_income = "B19025",
  fam_income = "B19101",
  poverty = "C17002",
  pov_age = "B17024",
  fam_poverty = "B17010",
  vehicle = "B25044",
  crowding = "B25014",
  housing_val = "B25075",
  mortgage = "B25091",
  rent = "B25074"
)
```

```{r}
fetch_tract <- table_nums %>%
	map(function(num) {
		get_acs(geography = "tract", table = num, year = year, state = "09", county = "01") %>%
			filter(GEOID %in% tracts) %>%
			inner_join(acs_vars, by = c("variable" = "name")) %>%
			inner_join(geo, by = c("GEOID" = "trct")) %>%
			mutate(level = "neighborhood")
	})
fetch_city <- table_nums %>%
	map(function(num) {
		get_acs(geography = "county subdivision", table = num, year = year, state = "09", county = "01") %>%
			mutate(name = str_extract(NAME, "^[\\w\\s]+(?= town,)")) %>%
			mutate(name2 = name) %>%
			filter(name %in% city_names) %>%
			inner_join(acs_vars, by = c("variable" = "name")) %>%
			mutate(level = "city")
	})
fetch_county <- table_nums %>%
	map(function(num) {
		get_acs(geography = "county", table = num, year = year, state = "09") %>%
			filter(GEOID %in% county) %>%
			inner_join(acs_vars, by = c("variable" = "name")) %>%
			mutate(name = str_extract(NAME, "^[\\w\\s]+(?=,)")) %>%
			mutate(name2 = name) %>%
			mutate(level = "county")
	})
fetch_state <- table_nums %>%
	map(function(num) {
		get_acs(geography = "state", table = num, year = year) %>%
			filter(GEOID == "09") %>%
			inner_join(acs_vars, by = c("variable" = "name")) %>%
			mutate(name = NAME) %>%
			mutate(name2 = name) %>%
			mutate(level = "state")
	})
fetch <- pmap(list(fetch_tract, fetch_city, fetch_county, fetch_state), bind_rows) %>%
	map(~replace_na(., list(share = 1.0)) %>% 
				mutate(wtd = estimate * share) %>%
				mutate(level = as.factor(level) %>% fct_relevel("neighborhood", "city", "county", "state")) %>%
				select(level, city, name, share, variable, label, wtd) %>%
				mutate(label = str_remove(label, "Total!!")) %>%
				mutate(label = factor(label, unique(label))) %>%
				group_by(level, city, name, label) %>%
				summarise(value = sum(wtd) %>% round()) %>%
				spread(key = label, value = value) %>%
				ungroup()
			)

table <- table_nums %>% map(~c())
```




Total population

```{r}
table$total_pop <- fetch$total_pop %>%
	rename(total_pop = Total) %>%
	gather(key = group, value = value, -level:-name) %>%
	mutate(share = NA %>% as.numeric())

table$total_pop
```

Sex by age: 0-5, 5-17, 18-24, 25-34, 35-44, 45-54, 55-64, 65+

```{r}
table$sex_age <- fetch$sex_age %>%
	mutate(
		under5 = add_grp(., 6, 30),
		age5_17 = add_grp(., 7:9, 31:33),
		age18_24 = add_grp(., 10:13, 34:37),
		age25_34 = add_grp(., 14:15, 38:39),
		age35_44 = add_grp(., 16:17, 40:41),
		age45_54 = add_grp(., 18:19, 42:43),
		age55_64 = add_grp(., 20:22, 44:46),
		age65plus = add_grp(., 23:28, 47:52)
	) %>%
	select(level:name, total_pop = Total, male = Male, female = Female, under5:age65plus) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("total_pop") %>%
	filter(group != "total_pop")

table$sex_age
```


Race: hispanic, non-hispanic white, black, asian, other

```{r}
# set name to total_pop_race to deal with duplicates
table$race <- fetch$race %>%
	mutate(other_race = add_grp(., 8, 10:12)) %>% 
	select(level:name, total_pop_race = Total, hispanic = 15, white = 6, black = 7, asian = 9, other_race) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("total_pop_race")

table$race
```


Family structure: total households; family households: married couple, single male, single female

```{r}
family <- fetch$family %>%
	select(level:name, total_hh = Total, family_hh = 5, married_hh = 6, male_hh = 8, female_hh = 9) %>%
	gather(key = group, value = value, -level:-name)

table$family <- bind_rows(
	family %>% filter(group %in% c("total_hh", "family_hh")) %>% make_share("total_hh"),
	family %>% filter(group != "total_hh") %>% make_share("family_hh") %>% mutate(group = factor(group) %>% fct_recode(family_hh_type = "family_hh"))
)

table$family
```


Children in families: married family with kid, male hh with kid, female hh with kid, as percentage of family households

```{r}
table$children <- fetch$children %>%
	select(level:name, family_hh = 4, married_hh_kids = 6, male_hh_kids = 13, female_hh_kids = 23) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("family_hh") %>%
	filter(group != "family_hh")

table$children
```


Vacancy & tenure: total housing units, vacant, occupied; owned & rented as percentage of occupied units

```{r}
vac_ten <- fetch$vacancy %>%
	select(level:name, vacant = Total) %>%
	inner_join(fetch$tenure, by = c("level", "city", "name")) %>%
	select(level:name, vacant, occupied = Total, owned = 6, rented = 7) %>%
	mutate(total_units = vacant + occupied) %>%
	gather(key = group, value = value, -level:-name) %>%
	mutate(group = as.factor(group) %>% fct_relevel("total_units", "vacant")) %>%
	arrange(group)

table$vacancy <- vac_ten %>% 
	filter(group %in% c("total_units", "vacant", "occupied")) %>% 
	make_share("total_units")
table$tenure <- vac_ten %>% 
	filter(group %in% c("occupied", "owned", "rented")) %>% 
	make_share("occupied") %>%
	filter(group != "occupied")

table$vacancy
table$tenure
```


disconnected youth: pop ages 16-19, share not working & not enrolled

```{r}
table$disconnect <- fetch$disconnect %>%
	mutate(disconnected = add_grp(., 13:14, 17:18, 27:28, 31:32)) %>%
	select(level:name, ages16_19 = Total, disconnected) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("ages16_19")

table$disconnect
```


```{r}
table$education <- fetch$education %>%
	mutate(less_than_hs = add_grp(., 5:12),
				 hs_plus = add_grp(., 13:20),
				 bach_plus = add_grp(., 17:20)
				 ) %>%
	select(level:name, age25plus = Total, less_than_hs:bach_plus) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("age25plus")

table$education
```

language + english speakers: pop age 5+; speak only english, non-english home language; non-english + speak less than very well; spanish + less than very well

```{r}
table$language2 <- fetch$language2 %>% 
	mutate(
		english_only = add_grp(., 6, 28, 50),
		other_home_lang = add_grp(., 7, 12, 17, 22, 29, 34, 39, 44, 51, 56, 61, 66),
		low_english = add_grp(., 9:11, 14:16, 19:21, 24:26, 31:33, 36:38, 41:43, 46:48, 53:55, 58:60, 63:65, 68:70),
		spanish = add_grp(., 7, 29, 51),
		spanish_low_eng = add_grp(., 9:11, 31:33, 53:55)
		) %>%
	select(level:name, age5plus = Total, english_only:spanish_low_eng) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("age5plus")

table$language2
```

employment: pop age 16+, labor force, civilian labor force, employed, unemployed, not in labor force. unemployment rate out of civ labor force

```{r}
labor <- fetch$labor %>%
	select(level:name, age16plus = Total, labor_force = 5, civ_labor_force = 6, civ_employed = 7, civ_unemployed = 8, not_in_force = 10) %>%
	gather(key = group, value = value, -level:-name)

table$labor <- bind_rows(
	labor %>% make_share("age16plus"),
	labor %>% 
		filter(group %in% c("civ_labor_force", "civ_unemployed")) %>% 
		make_share("civ_labor_force") %>% 
		mutate(group = as.factor(group) %>% fct_recode(unemployment = "civ_unemployed")) %>%
		filter(group != "civ_labor_force")
)

table$labor
```

commuting: workers age 16+, work at home, don't work at home, car alone, carpool, public transit, walk + bike, cab/motorcycle/other

```{r}
table$commute <- fetch$commute %>%
	mutate(
		work_not_home = add_grp(., 5, 13, 19:23),
		walk_bike = add_grp(., 21:22),
		other_transit = add_grp(., 19, 20, 23)
	) %>% 
	select(level:name, workers = Total, work_home = 24, work_not_home, drive_alone = 6, carpool = 7, public_transit = 13, walk_bike, other_transit) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("workers")

table$commute
```

working parents: number of kids in families; kids in 2-parent family both working + kids in 1-parent family with parent working

```{r}
table$parent_work <- fetch$parent_work %>%
	mutate(kids_all_parents_work = add_grp(., 7, 13, 16, 20, 26, 29)) %>%
	select(level:name, kids_in_fams = Total, kids_all_parents_work) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("kids_in_fams")

table$parent_work
```

occupation: employed; management, professional, related; service; sales + office; natural resources, construction, maintenance; production, transport, material moving

```{r}
table$occupation <- fetch$occupation %>%
	mutate(
		management = add_grp(., 6, 42),
		service = add_grp(., 22, 58),
		sales = add_grp(., 30, 66),
		resources = add_grp(., 33, 69),
		production = add_grp(., 37, 73)
	) %>%
	select(level:name, employed = Total, management:production) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("employed") %>%
	mutate(group = factor(group) %>% fct_recode(employed_occ = "employed"))

table$occupation
```

household income: total households; < 10k; 10k-14.9; 15k-24.9; < 25k; 25k-49.9; < 50k; 50k-74.9; 75k-99.9; > 100k; 100k-149.9; 150k-199.9; > 200k

```{r}
table$hh_income <- fetch$hh_income %>%
	mutate(
		hh_inc15_25 = add_grp(., 7:8),
		hh_inc_under25 = add_grp(., 5:8),
		hh_inc25_50 = add_grp(., 9:13),
		hh_inc_under50 = add_grp(., 5:13),
		hh_inc50_75 = add_grp(., 14:15),
		hh_inc_100plus = add_grp(., 17:20),
		hh_inc100_150 = add_grp(., 17:18)
	) %>%
	select(level:name, households = Total, hh_inc_under10 = 5, hh_inc10_15 = 6, hh_inc15_25, hh_inc_under25, hh_inc25_50, hh_inc_under50, hh_inc50_75, hh_inc75_100 = 16, hh_inc_100plus, hh_inc100_150, hh_inc150_200 = 19, hh_inc_200plus = 20) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("households") %>%
	mutate(group = factor(group) %>% fct_recode(households_inc = "households"))

table$hh_income
```

average household income: divide agg income by number of households

```{r}
table$agg_income <- fetch$agg_income %>%
	inner_join(fetch$hh_income %>% select(level:name, Total), by = c("level", "city", "name")) %>%
	select(level:name, agg = 4, households = 5) %>%
	mutate(avg_hh_inc = round(agg / households)) %>%
	select(-agg:-households) %>%
	gather(key = group, value = value, -level:-name) %>%
	mutate(share = NA %>% as.numeric())

table$agg_income
```

family income: same groups as household

```{r}
table$fam_income <- fetch$fam_income %>%
	mutate(
		fam_inc15_25 = add_grp(., 7:8),
		fam_inc_under25 = add_grp(., 5:8),
		fam_inc25_50 = add_grp(., 9:13),
		fam_inc_under50 = add_grp(., 5:13),
		fam_inc50_75 = add_grp(., 14:15),
		fam_inc_100plus = add_grp(., 17:20),
		fam_inc100_150 = add_grp(., 17:18)
	) %>%
	select(level:name, families = Total, fam_inc_under10 = 5, fam_inc10_15 = 6, fam_inc15_25, fam_inc_under25, fam_inc25_50, fam_inc_under50, fam_inc50_75, fam_inc75_100 = 16, fam_inc_100plus, fam_inc100_150, fam_inc150_200 = 19, fam_inc_200plus = 20) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("families") %>%
	mutate(group = factor(group) %>% fct_recode(families_inc = "families"))

table$fam_income
```

ratio to poverty level: number with status; number at 1xfpl, 2x, > 2x

```{r}
table$poverty <- fetch$poverty %>%
	mutate(
		poverty = add_grp(., 5:6),
		low_inc = add_grp(., 5:10)
	) %>%
	select(level:name, pov_determined = Total, poverty, low_inc, twice_fpl = 11) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("pov_determined")

table$poverty
```

families in poverty: poverty determined, number under fpl

```{r}
table$fam_poverty <- fetch$fam_poverty %>%
	select(level:name, families = 4, family_poverty = 5) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("families") %>%
	mutate(group = factor(group) %>% fct_recode(families_pov = "families"))

table$fam_poverty
```

vehicles: occupied housing units, 0, 1, 2+ vehicles

```{r}
table$vehicle <- fetch$vehicle %>%
	mutate(
		hh_0_vehicle = add_grp(., 6, 13),
		hh_1_vehicle = add_grp(., 7, 14),
		hh_2plus_vehicle = add_grp(., 8:11, 15:18)
	) %>%
	select(level:name, occupied_units = Total, hh_0_vehicle:hh_2plus_vehicle) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("occupied_units")

table$vehicle
```

overcrowding: 1 or less occupants; 1.01 or more

```{r}
table$crowding <- fetch$crowding %>%
	mutate(
		not_crowded = add_grp(., 6:7, 12:13),
		overcrowded = add_grp(., 8:10, 14:16)
	) %>%
	select(level:name, households = Total, not_crowded, overcrowded) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("households") %>%
	filter(group != "households")

table$crowding
```

housing value:  owner-occupied units; house values < 50k; 50k-99.9; 100k-149.9; 150k-199.9; > 200k; 200k-249.9; 250k-299.9; > 300k

```{r}
table$housing_val <- fetch$housing_val %>%
	mutate(
		val_under50 = add_grp(., 5:12),
		val50_100 = add_grp(., 13:17),
		val100_150 = add_grp(., 18:19),
		val150_200 = add_grp(., 20:21),
		val_200plus = add_grp(., 22:30),
		val_300plus = add_grp(., 24:30)
	) %>%
	select(level:name, owned_units = Total, val_under50, val50_100, val100_150, val150_200, val_200plus, val200_250 = 22, val250_300 = 23, val_300plus) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("owned_units") %>%
	mutate(group = factor(group) %>% fct_recode(owned_units_val = "owned_units"))

table$housing_val
```

housing cost burden, mortgage + rent; owned units, owned burden, owned severe burden; rented units, rented burden, rented severe burden; both

```{r}
mortgage <- fetch$mortgage %>%
	mutate(
		owned_cost30 = add_grp(., 11:14, 22:25),
		owned_cost50 = add_grp(., 14, 25)
	) %>%
	select(level:name, owned_units = Total, owned_cost30, owned_cost50)

rent <- fetch$rent %>%
	mutate(
		rent_cost30 = add_grp(., 9:12, 18:21, 27:30, 36:39, 45:48, 54:57, 63:66),
		rent_cost50 = add_grp(., 12, 21, 30, 39, 48, 57, 66)
	) %>%
	select(level:name, rented_units = Total, rent_cost30, rent_cost50)

cost <- inner_join(mortgage, rent, by = c("level", "city", "name")) %>%
	mutate(units = owned_units + rented_units) %>%
	mutate(units_cost30 = owned_cost30 + rent_cost30, units_cost50 = owned_cost50 + rent_cost50) %>%
	gather(key = group, value = value, -level:-name)

table$cost <- bind_rows(
	cost %>% filter(str_detect(group, "^owned")) %>% make_share("owned_units"),
	cost %>% filter(str_detect(group, "^rent"))  %>% make_share("rented_units"),
	cost %>% filter(str_detect(group, "^units")) %>% make_share("units")
) %>%
	mutate(group = factor(group) %>% fct_recode(owned_units_burden = "owned_units", rented_units_burden = "rented_units", units_burden = "units"))
table$mortgage <- NULL
table$rent <- NULL

table$cost
```

immigrants: total pop, total foreign born; entered 2010 or later, naturalized, not citizen

```{r}
table$immigration <- fetch$immigration %>%
	mutate(
		foreign_born = add_grp(., 7, 12, 17, 22),
		naturalized = add_grp(., 8, 13, 18, 23),
		not_citizen = add_grp(., 9, 14, 19, 24)
	) %>%
	select(level:name, total_pop_fb = Total, foreign_born, entered_2010later = 5, naturalized, not_citizen) %>%
	gather(key = group, value = value, -level:-name) %>%
	make_share("total_pop_fb")

table$immigration
```

poverty by age: poverty determined, under fpl: under 6yrs, under 18yrs, 65+

```{r}
pov_age <- fetch$pov_age %>%
	mutate(
		under6_status = add_grp(., 5),
		under6_poverty = add_grp(., 6:8),
		under18_status = add_grp(., 5, 18, 31),
		under18_poverty = add_grp(., 6:8, 19:21, 32:34),
		over65_status = add_grp(., 109, 122),
		over65_poverty = add_grp(., 110:112, 123:125)
	) %>%
	select(level:name, under6_status:over65_poverty) %>%
	gather(key = group, value = value, -level:-name)

table$pov_age <- bind_rows(
	pov_age %>% filter(str_detect(group, "^under6")) %>% make_share("under6_status"),
	pov_age %>% filter(str_detect(group, "^under18")) %>% make_share("under18_status"),
	pov_age %>% filter(str_detect(group, "^over65")) %>% make_share("over65_status")
)

table$pov_age
```

bind together

```{r}
rearrange_cols <- function(.data, omit = 1) {
	x <- c(1:ncol(.data))[-omit]
  half <- length(x) / 2
  vals <- x[1:half]
  shares <- x[(half + 1):length(x)]
  # transpose: works like zip in python, returns vector of positions with est & moe interwoven
  # names(.data)[shares] <- names(.data)[vals]
  arranged <- list(vals, shares) %>% transpose() %>% unlist()
  return(.data %>% select(omit, arranged))
}
```


```{r}
bound <- table %>% bind_rows() %>% mutate(group = factor(group, unique(group)))

df <- inner_join(
	bound %>% select(-share) %>% spread(key = group, value = value),
	bound %>% select(-value) %>% spread(key = group, value = share),
	by = c("level", "city", "name"), suffix = c("_num", "_per")
) %>% 
	rearrange_cols(omit = 1:3)

write_csv(df, sprintf("%s BPT Stamford extended.csv", year))
```

